### 2024-09-17 base-rates.qmd
20240917 - clarifications

### Example 2: Does a Woman Have Cancer? {#sec-nateSilverExample2}

@@

- the [prior probability](#sec-baseRate) that she has cancer: 1.4%
- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she has cancer: 75%
- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she does *not* have cancer: 10%

@@

Fantasy performance—and behavior more generally—is challenging to predict.
People commonly demonstrate [biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies) when making predictions.
People tend to ignore base rates ([base rate fallacy](#sec-fallaciesBaseRate)) when making predictions.
They also tend to confuse inverse conditional probabilities ([conditional probability fallacy](#sec-inverseFallacy)).
[Bayes' theorem](#sec-bayesTheorem) provides a way to convert from one conditional probability to its inverse conditional probability using the [base rate](#sec-baseRate) of each event.
There are various ways to account for [base rates](#sec-baseRate) for more accurate predictions, including through the use of [actuarial formulas](#sec-accountForBaseRatesActuarial) and [Bayesian updating](#sec-bayesianUpdating).
[Bayesian updating](#sec-bayesianUpdating) uses [Bayes' theorem](#sec-bayesTheorem) to calculate a posttest probability from a [pretest probability](#sec-baseRate) and a test result ([likelihood ratio](#sec-diagnosticLikelihoodRatio)).
The [probability nomogram](#sec-probabilityNomogram) is a visual approach to [Bayesian updating](#sec-bayesianUpdating).

>>>>

### Example 2: Does a Person Have Breast Cancer? {#sec-nateSilverExample2}

@@

- the [prior probability](#sec-baseRate) that she has breast cancer: 1.4%
- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she has breast cancer: 75%
- the [conditional probability](#sec-conditionalProbability) that she has a positive test given that she does *not* have breast cancer: 10%

@@

Fantasy performance—and behavior more generally—is challenging to predict.
People commonly demonstrate [biases](#sec-cognitiveBiases) and [fallacies](#sec-fallacies) when making predictions.
People tend to ignore base rates ([base rate fallacy](#sec-fallaciesBaseRate)) when making predictions.
They also tend to confuse inverse conditional probabilities ([conditional probability fallacy](#sec-inverseFallacy)).
[Bayes' theorem](#sec-bayesTheorem) provides a way to convert from one conditional probability to its inverse conditional probability using the [base rate](#sec-baseRate) of each event.
There are various ways to account for [base rates](#sec-baseRate) for more accurate predictions, including through the use of [actuarial formulas](#sec-accountForBaseRatesActuarial), [Bayesian updating](#sec-bayesianUpdating), and more [informal approaches](#sec-informalUpdating).
[Bayesian updating](#sec-bayesianUpdating) uses [Bayes' theorem](#sec-bayesTheorem) to calculate a posttest probability from a [pretest probability](#sec-baseRate) and a test result ([likelihood ratio](#sec-diagnosticLikelihoodRatio)).
The [probability nomogram](#sec-probabilityNomogram) is a visual approach to [Bayesian updating](#sec-bayesianUpdating).
